---
title: "Envs 193DS week 7 coding workshop"
format: html
editor: visual
execute: 
  warning: FALSE
  message: FALSE 
---

# set up 

```{r libraries}
library(tidyverse)
library(here) 
library(lterdatasampler)

library(performance)
library(broom) #puts all outputs from model into a table
library(flextable) #whole manual online, allows you to create tables that render nicely 
library(ggeffects) #get predictions from models and plot them...
library(car) #pull out ANOVA tables specifically for linear models 
library(naniar)
```

# Linear models 

How does stem length _predict_ stem dry mass? 

```{r filtering-data}
maples_data <- hbr_maples |> 
                filter(year == 2003 & watershed == "Reference")

```

Visualizing missing data 
```{r missing-data-var}
gg_miss_var(maples_data) #we have no missing observations for filtered data frame 

```

Create exploratory data visualization: 

```{r explore-data-viz}
ggplot(data = maples_data, aes(x = stem_length, y = stem_dry_mass)) + 
  geom_point() #just looking at the plot, would hypothesize that there is a positive relationship

```

#Let's try a model: 

```{r linear-model-maples}

#lm() is how you run a linear model 
#first, write model -- then check assumptions -- need to do this before interpreting model! 
maples_model <- lm(stem_dry_mass ~ stem_length, data = maples_data)

maples_model

                   

```
Check assumptions:
1. linear relationship between variables: yes! (used exploratory data viz)
2. independence of errors: yes! (making that assumption based on how the data were collected)
3. homoskedasticity of errors: yes! (making deicision from residuals vs fitted/scale-location plots)
4. normally distributed errors: yes! (looking at qqplot of rediduals)

```{r checking-assumptions}
#TRY running plot(maples_model) in console 
#this will give diagnostic plots 
#residuals vs fitted seems homoscedastic 
#normal q-q looks pretty normal 
#scale-location -- looks right, line is pretty straight
#residuals vs leverage -- looking for labeled points, there are some but they probs dont influnce predictions much because are not outside of dotted line range -> does not seem that there are outliers affecting model predictions 

#display all on same grid 
#set graphical parameters 
par(mfrow = c(2,2))
plot(maples_model)
#we do not want the same dimensions for every plot 
#so, turn of 2x2 grid

```


```{r turning-off-the-grid, results = FALSE}
dev.off()
```

